Cost- and Deadline-Constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-25 paper/paper.tex
Update conclusions

In this paper we addressed the interesting and important new problem of
scheduling and resource provisioning for scientific workflow ensembles
on IaaS clouds. The problem is different from previous work on grid and
utility grid scheduling in that cloud infrastructures provide more control
over the resources, so the resource provisioning plan can be adjusted
according to the requirements of the application. Therefore the problem
space becomes larger: it requires not only an efficient mapping of tasks
to available resources, but also the selection of the best resource
provisioning plan. Formulating the problem as a maximization of the
number of prioritized workflows completed from the ensemble is also novel and  
requires workflows to be admitted or rejected based on their estimated
resource demands. We believe that this bi-constrained problem is highly
relevant because such constraints are commonly imposed on many real-world
projects. The approach is also directly applicable to grid environments
that provide resource reservations and charge service units for resource use.

We have considered both static and dynamic scheduling approaches and
developed the DPDS, WA-DPDS and SPSS algorithms based on these
strategies. The algorithms were evaluated on ensembles of synthetic
workflows, which were generated based on real scientific applications. For
the purposes of evaluation we have developed a simulator that models: the
cloud infrastructure, and a workflow engine with tightly-coupled scheduler
and provisioner modules. The results of our simulation studies indicate
that the two algorithms that take into account information about the
structure of the workflow and task runtime estimates (WA-DPDS and SPSS)
yield better results than the simple on-line priority-based scheduling
strategy (DPDS).

During our work, several interesting lessons were learned from simulating the execution of
workflow ensembles on cloud infrastructures. One result is that the simple
strategy used by DPDS performs relatively well in terms of the number of
workflows completed and the high resource utilization it can achieve. However,
some of these workflows may be of lower priority, and the high utilization
may be achieved by running tasks from workflows that cannot be completed
given the constraints. On the other hand, DPDS may be the only choice when
estimates of task runtimes are not available in advance. We also observed
that the WA-DPDS and SPSS algorithms are able to find better quality
solutions in terms of the target metrics: this is primarily a result of the
admission algorithms, which reject workflows that are too large to be
completed given the constraints, and accept more smaller workflows that
can be completed.

Another lesson learned from our studies is the effect of task granularity on
workflow ensembles. The Montage and CyberShake workflows have fine-grained tasks
with execution times on the order of seconds to minutes, while LIGO, Epigenomics
and SIPHT have coarser-grained tasks with runtimes ranging from several minutes
to hours. We observed that the dynamic algorithms achieved better results on the
fine grained workflows since the simple random scheduling strategy leads to good
load-balancing and high resource utilization for these ensembles. On the other
hand, when tasks are longer, the dynamic algorithms tend to use up the budget by
running tasks from low-priority workflows while the more sophisticated static
scheduling strategy of SPSS is more circumspect. In the future we plan to
investigate further the effects of task granularity on ensemble scheduling in a more
systematic fashion.

Our current study suggests many areas for future work. One interesting
issue that we plan to investigate is how the static and dynamic algorithms
perform with uncertain task runtimes and failures, which are known to have
an important influence on overall system performance~\cite{Sakellariou2010,Dongarra2007}.
We also plan to extend our application and infrastructure model to explicitly
include the various data storage options available on clouds (right now we only
model data access as part of the task time and do not include data costs). A previous
experimental study on that subject~\cite{Juve2010} suggests that the data
demands of scientific workflows have a large impact on not only the
execution time, but also on the cost of workflows in commercial clouds.
Finally, we plan to investigate heterogeneous environments that include
multiple VM types and cloud providers, including private and community clouds,
which will make the problem even more complex and challenging~\cite{Marshall2010, Vockler2011, Juve2010}.

We believe that the results of our study can be applied in practice to develop
tools that assist researchers in planning their large-scale computational
experiments. The estimates of costs, runtimes and number of workflows completed
that can be obtained from both the static algorithms and from the simulation
runs, constitute valuable hints for the planning decisions and assessing the
associated trade-offs. For this purpose our future work includes also further
extension of our experimental studies~\cite{Juve2009, Vockler2011} of workflows on
available cloud infrastructures to develop more accurate performance and cost
models of the applications. This should lead to better understanding of the
problems and costs of doing science in the cloud.

@@

% ACM style
%\bibliographystyle{abbrv}
% IEEE style
\bibliographystyle{IEEEtran}

>>>>

In this paper we addressed the interesting and important new problem of scheduling and resource provisioning for scientific workflow ensembles on IaaS clouds. The problem is different from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over the resources, so the resource provisioning plan can be adjusted according to the requirements of the application. Therefore the problem space becomes larger; it requires not only an efficient mapping of tasks to available resources, but also the selection of the best resource provisioning plan. Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands. We believe that this bi-constrained problem is highly relevant because such constraints are commonly imposed on many real-world projects. The approach is also directly applicable to grid environments that provide resource reservations and charge service units for resource use.

The DPDS, WA-DPDS and SPSS algorithms were developed based on both static and dynamic scheduling approaches. The algorithms were evaluated on ensembles of synthetic workflows, which were generated based on real scientific applications. For the purposes of evaluation we have developed a simulator that models the cloud infrastructure, and a workflow engine with tightly-coupled scheduler and provisioner modules.

The results of our simulation studies indicate that the two algorithms that take into account information about the structure of the workflow and task runtime estimates (WA-DPDS and SPSS) yield better results than the simple priority-based scheduling strategy (DPDS), which makes provisioning decisions purely based on resource utilization. On the other hand, DPDS may be the only choice when estimates of task runtimes are not available in advance.

In cases where there are no provisioning delays, task runtime estimates are good, and failures are rare, we found that the static algorithm performs significantly better than the dynamic algorithms. However, when runtime estimates are poor, delays are present, or failures are common, the plans produced by the static algorithm are obstructed and it frequently exceeds the budget and deadline constraints. In comparison, the dynamic algorithms are able to quickly adapt to even verly large delays, poor estimates, and failures to prevent the ensemble from exceeding the constraints.

In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms. For wide, fine-grained workflows, such as CyberShake and Montage, however, the dynamic algorithms frequently produce performance that is as good or better than SPSS,  because they are better able to pack the smaller tasks onto idle VMs close to the deadline than SPSS, which distributes deadlines to tasks in a way that prevents them from starting late.

For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive due to its $O(n^2)$ algorithmic complexity. Our simulation results showed that ensembles of 100 workflows with 800 tasks often took 10 minutes or more to plan. This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan. In comparison, the $O(n)$ complexity of the dynamic algorithms makes them much more attractive for large scale problems.

This study suggests many areas for future work. It would be interesting to extend the application and infrastructure model to explicitly include the various data storage options available on clouds (right now data access is modeled as part of the task execution time and data costs are not considered). A previous experimental study \cite{Juve2010} suggests that the data demands of scientific workflows have a large impact on not only the execution time, but also on the cost of workflows in commercial clouds. Finally, it would be interesting to investigate heterogeneous environments that include multiple VM types and cloud providers, including private and community clouds, which will make the problem even more complex and challenging~\cite{Marshall2010, Vockler2011, Juve2010}.

The results of this study can also be applied to develop tools that assist researchers in planning their large-scale computational experiments. The estimates of cost, runtime, and number of workflows completed that can be obtained from both the static algorithms and from the simulation runs, constitute valuable hints for planning ensembles and evaluating the associated trade-offs.

@@

% ACM style
%\bibliographystyle{abbrv}

% IEEE style
\bibliographystyle{IEEEtran}

% This balances the columns by inserting a column break at the given reference
\IEEEtriggeratref{13}
